Analysing Constraint Grammars with a SAT-solver

### 2015-10-24 paper/lrec/problem_desc.tex
remove text

Constraint Grammar (CG, \cite{karlsson1995constraint})
% \citep[CG][]{karlsson1995constraint} 
is a formalism used to disambiguate morphologically analysed text. 
A grammar consists of rules that target specific analyses for selection or removal, based on contextual tests. For example, the following rule
\begin{itemize}
\item[] \texttt{REMOVE verb IF (-1 det) ;}
\end{itemize}
removes a verb reading from a word which is preceded by adeterminer.

Given the following text,
\begin{itemize}
\item[] 
\begin{verbatim}
"<the>"
        "the" det def
"<bear>"
        "bear" noun sg
        "bear" verb pl
\end{verbatim}
\end{itemize}
the rule will match to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
However, if the target word has only one remaining analysis, then the rule will not apply, even if the condition is met.

@@

As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
%\cite{voutilainen2004} gives a detailed account about best practices of grammar writing and efficient use of hand-annotated corpora to aid the grammar development.
%\cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
%Our tools complement the existing methods in CG analysis and optimisation.
%While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
%Authors who have a large hand-annotated corpus can also benefit from our tools.
With our tools, the users can diagnose and improve their grammars.
They can ask specific questions about a given grammar, such as:

\begin{itemize}
\item Are there rules that contradict each other?
\item Are there rules that will never fire?
\item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$
%\item Generate a sequence that is ambiguous but doesn't trigger any rules
\end{itemize}

Some examples where a rule contradicts another are listed below. R1
can also be a set of rules: for instance, if one rule removes a verb in
context $C$, and another in context $\neg C$, together these rules
remove a verb in all possible cases, disabling any future rule that
targets verbs.
\begin{itemize}
\item If a rule appears twice, the second occurrence will be disabled by the first
\item $R1$ selects something in a context, $R2$ removes it
\item $R1$ removes something from $R2$'s context, so $R2$ can never apply
\end{itemize}

Our technique requires an existing morphological lexicon, but no corpus. 
The lexicon is needed in order to generate realistic ambiguities: e.g. a word can be ambiguous between noun and verb, but not between adjective and punctuation. 
Constraints for the structure of the sequences come from the CG rules themselves.
We encode both of these constraints as a satisfiability problem, and use a SAT-solver to generate answers to the previous questions.

>>>>

Constraint Grammar (CG, \cite{karlsson1995constraint})
% \citep[CG][]{karlsson1995constraint} 
is a formalism used to disambiguate morphologically analysed text. 
A grammar consists of rules that target specific analyses for selection or removal, based on contextual tests. For example, the following rule
\begin{itemize}
\item[] \texttt{REMOVE verb IF (-1 det) ;}
\end{itemize}
removes a verb reading from a word which is preceded by a determiner.
Given the following text,
\begin{itemize}
\item[] 
\begin{verbatim}
"<the>"
        "the" det def
"<bear>"
        "bear" noun sg
        "bear" verb pl
\end{verbatim}
\end{itemize}
the rule will match to the word \emph{bear}, and remove the analysis \texttt{verb pl}.
However, if the target word has only one remaining analysis, then the rule will not apply, even if the condition is met.

@@

As CGs grow larger, it gets harder for the grammar writers and users
to keep track of all the rules and their interaction.
With our tools, the users can diagnose and improve their grammars.
Some examples of conflicts are listed below:
\begin{itemize}
\item If a rule appears twice, the second occurrence will be disabled by the first
\item $R$ selects something in a context, $R'$ removes it
\item $R$ removes something from the context of $R'$, so $R'$ can never
  apply
\item $R$ has an internal conflict, such as non-existent
tag combination, or contradicting requirements for a context word
\end{itemize}
$R$ can also be a set of rules: for instance, if one rule removes a verb in
context $C$, and another in context $\neg C$, together these rules
remove a verb in all possible cases, disabling any future rule that
targets verbs.

%In addition to analysing the whole grammar,
Our method can be used to
test out individual rules. For instance, we can ask the tool to
generate example sequences that trigger rule(s) $R$ but not rule(s) $R'$. 
This functionality can help the grammar writer to see
if they have implemented the rules in the intended way.

% They can ask specific questions about a given grammar, such as:
% \begin{itemize}
% \item Are there rules that contradict each other?
% \item Are there rules that will never fire?
% \item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$
% %\item Generate a sequence that is ambiguous but doesn't trigger any rules
% \end{itemize}
% Our technique requires an existing morphological lexicon, compatible
% with the tag set used in the grammar, but no corpus. 
% The lexicon is needed in order to limit the possible tag combinations.
% Constraints for the structure of the sequences come from the CG rules themselves.
% We encode both of these constraints as a satisfiability problem, and use a SAT-solver to generate answers to the previous questions.
